---
title: "JSC370 Final Project"
author: "Yufan Qian"
output: 
    html_document:
        toc: TRUE
        toc_float: TRUE
---

### Introduction

Nowadays, we have so many apps on our phone and computers and everyone's life is associated with them. The project is focused on factors affecting people's opinion on different apps and whether they favor some apps with particular characteristics than others. I have narrowed my target to apps on Google Play Store.

The primary question of interest is: What are the factors for different app's rating?

The variables we are interested in:

- `Rating`: the rating of the app. We want to find all the factors affecting the rating.
- `Reviews`: the number of reviews of the app. Could indicate the popularity of the app.
- `Category`: the category of the app. Examples are: `art and design`, `Game`, `Tools`, etc.
- `Size`: the size of the app. This is a new variable compared with my midterm project, and we will discuss about it below.
- `Installs`: number of installs of the app.

Download my report here: https://github.com/lynnqian/JSC370_final_project/blob/master/JSC%20Final%20files%20/JSC_final.pdf

### Interactive Visualizations

```{r message=FALSE, echo=FALSE, results='hide'}
# load the library we need
library(data.table)
library(dplyr)
library(tidyr)
library(ggplot2)
library(mgcv)
library(plotly)

knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE) 

# load in the data
playstore_raw <- data.table::fread("data/googleplaystore.csv",fill=TRUE)

# all the variables with missing values
names(which(sapply(playstore_raw, anyNA)))

# remove observations with missing values for Rating
playstore <- dplyr::filter(playstore_raw, !is.na(Rating))
# keep only one observation for each apps
playstore <- playstore[match(unique(playstore$App), playstore$App),]

dim(playstore)
names(which(sapply(playstore, anyNA)))

playstore <- playstore[Rating <= 5]

# convert from type character to type integer
playstore$Reviews <- as.numeric(playstore$Reviews)

# remove the last character: +
playstore$Installs <- substr(playstore$Installs, 1, nchar(playstore$Installs)-1)
# convert from type character to type integer
playstore$Installs <- as.numeric(gsub(",","",playstore$Installs))

playstore <- playstore[!(playstore$Size == "Varies with device")]
playstore$Size <- substr(playstore$Size, 1, nchar(playstore$Size)-1)
playstore$Size <- as.numeric(playstore$Size)
playstore$Size <- playstore$Size*1000000

playstore$review_level <- ifelse(playstore$Reviews < 1000, "almost no reviews", 
                                    ifelse(playstore$Reviews < 10000, "medium reviews", 
                                           ifelse(playstore$Reviews < 50000, "many reviews", "lots of reviews")))

playstore$install_level <- ifelse(playstore$Installs < 1000, "not popular", 
                                    ifelse(playstore$Installs < 10000, "normal", 
                                           ifelse(playstore$Installs < 50000, "popular", "super popular")))
```

#### Installs vs Reviews

Here, we want to see if the two variables: `Installs` and `Reviews` can both be used to represent popularity of the app by comparing them using the following line graph.

```{r}
playstore %>%
  filter(Category == "GAME") %>%
  plot_ly(x=~Rating, y=~Installs/30, type = "scatter", name = "Installs", mode = "lines") %>%
  add_lines(x=~Rating, y=~Reviews, type = "scatter", name = "Reviews", mode = "lines")
```

#### Line graph

This is a line plot with all the categories of the apps involved and combined with `Reviews`, and we want to see if there is any patterns between `Category` of the app and the its `Rating`.

```{r}
line_plot <- ggplot(playstore, aes(x = Reviews, y = Rating, colour = Category)) + 
  geom_line()

ggplotly(line_plot)
```


#### Size graph

Here, we added a new variable `Size` to see whether the `Size` of an app would affect its `Rating`.

```{r}
# change the type of the size
playstore$new_size <- playstore$Size/1000000
playstore %>%
  plot_ly(x = ~Reviews, y = ~Rating, type = 'scatter', mode = "markers", color = ~Category,  marker = list(size=~new_size/10, sizemode = 'diameter', opacity = 0.5))
```


In order to see the patterns more clearly, I created a `new_playstore` dataset with only apps that has reviews < 300,000 so that we could zoom in the left hand side of the graph.

```{r}
new_playstore <- playstore[playstore$Reviews < 300000]
new_playstore %>%
  plot_ly(x = ~Reviews, y = ~Rating, type = 'scatter', mode = "markers", color = ~Category,  marker = list(size=~new_size/10, sizemode = 'diameter', opacity = 0.5))
```



